### 2013-07-03 mcmc.tex
section headings
sample the posterior; don't sample the likelihood

...You can't sample a likelihood!  You can only sample a posterior
pdf!..
why not?..
Give a sense of the difficulties you get into if you fail here; like
discussion with Sven Kreiss on 2013-02-14...

@@

\section{stretch move and other variants}

>>>>

MCMC is used to obtain samples $x_k$ from PDF $p(x)$,
  or a badly normalized all-positive function $f(x)$
  that is different from the PDF by an unknown factor $Z$.
In the context of data analysis,
  MCMC is usually being used to obtain samples
  that are \emph{parameter values} for a probabilistic model of some data.

That is, for the purposes of this \documentname, we will imagine that
  there is some (possibly large block of) data $D$,
  and there are parameters $\theta$ of a \emph{likelihood function}
  $p(D\given\theta)$.%
\note{...something here about the definition of a model, for our purposes...}
  The likelihood function is used to compute the probability (density) at any
  particular observed data set, given any parameter setting.%
\note{...strictly this is only correctly called the ``likelihood'' when the data are treated as fixed...}
In standard probabilistic inference,
  the likelihood is used to update beliefs about the parameters.%
\note{...something here about us being Bayesian;
   if you aren't Bayesian, why are you considering MCMC for your data analysis?..
   probabilities are representations of beliefs not frequencies...
   beliefs can be thought of as subjective or ``official''...
   etc...}
In particular, if you started with prior probabilistic beliefs
  represented by a prior PDF $p(\theta)$,
  your posterior beliefs can be represented by a posterior PDF
  $p(\theta\given D)$, created via Bayes's rule:
\begin{eqnarray}
p(\theta\given D) = \frac{1}{Z}\,p(D\given\theta)\,p(\theta)
\quad,
\end{eqnarray}
  where $Z$ is a normalization constant.
MCMC is used in the context of data analysis
  (and hence the context of this \documentname)
  to sample posterior PDFs.

Importantly, \emph{you can't use MCMC to draw samples $\theta_k$ from the likelihood $p(D\given\theta)$!}
Why not?
Because $p(D\given\theta)$ is a PDF for the \emph{data} $D$ not a PDF for the \emph{parameters} $\theta$.
Many times data analysts believe they are sampling the likelihood,
  when in fact they are sampling a posterior PDF but with implicit priors.

...Give a sense of the difficulties you get into if you fail here; like
discussion with Sven Kreiss on 2013-02-14...

...Explain how your code \emph{should look} with a prior function and a likelihood function...

@@

\section{Ensembles, Hamiltonians, and other variants}
